MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex
top or main eigenvector, consistency!
major figure rework

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[h]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see section~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see section~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see section ~\ref{sec:vis}. Section~\ref{sec:results} describes our experimental results, Section~\ref{sec:param_choices} presents our parameter choices and  Section~\ref{sec:user_eval} summarizes our user evaluations.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=0.5\textwidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}a shows a volume rendering of the dataset. Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis. Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=0.5\textwidth]{images_pvis/reliable_hessian.pdf}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
\label{fig:reliable_hessian}
\end{figure}

@@

\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=0.3\textwidth]{imagesMT2014/image4}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on such assumptions on our data as local orientation and connectivity (Section~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb] 
  \centering  	
  \includegraphics[width=0.5\textwidth]{images_pvis/figure6.png}
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (C) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context.
  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\includegraphics[width=0.4\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}

>>>>

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.475\textwidth]{images_pvis/workflow.pdf}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see section~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see section~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see section ~\ref{sec:vis}. Section~\ref{sec:results} describes our experimental results, Section~\ref{sec:param_choices} presents our parameter choices and  Section~\ref{sec:user_eval} summarizes our user evaluations.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=\linewidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}a shows a volume rendering of the dataset. Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis. Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=\linewidth]{images_pvis/reliable_hessian.pdf}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
\label{fig:reliable_hessian}
\end{figure}

@@

\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=\linewidth]{images_pvis/mt_gen.pdf}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
Based on such assumptions on our data as local orientation and connectivity (Section~\ref{sec:char_data}), the order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb] 
  \centering  	
  \includegraphics[width=\linewidth]{images_pvis/clustering.pdf}
  %\includegraphics[width=0.45\textwidth]{imagesMT2014/image_clustering}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (C) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context.
  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\includegraphics[width=\linewidth]{images_pvis/dataset2.pdf}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}
